---
title: "COVID-19 Canada"
output: 
  flexdashboard::flex_dashboard:
    css: style.css
    social: menu
    # source_code: embed
    orientation: rows
    vertical_layout: scroll
    theme: flatly
    includes:
      after_body: footer.html
    runtime: shiny
---

```{r setup, echo=FALSE,results='hide',message = FALSE, warning = FALSE}
#------------------ Packages ------------------
library(flexdashboard) #dashboard;
library(shiny) #shiny app;
library(dplyr) #data tool;
library(RCurl) #read data url;
library(plotly) #visual tool;

options(scipen = 999)
# install_github("ishaberry/Covid19Canada") #COVID-19; Canada Open Data ishaberry/Covid19Canada
# can't read the working group data from repo directly;

#------------------ Read Canada data ------------------
#read raw github data;
# x1 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/cases.csv")
# x2 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/mortality.csv")
x1 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_prov/cases_timeseries_prov.csv") #new time series data, apr13;
x2 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_prov/mortality_timeseries_prov.csv") #new time series data, apr13;
x3 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/recovered_cumulative.csv")
x4 <- getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/testing_cumulative.csv")

can_c <- read.csv(text=x1, header = TRUE, sep = ",", encoding = 'UTF-8')
can_d <- read.csv(text=x2, header = TRUE, sep = ",", encoding = 'UTF-8')
can_r <- read.csv(text=x3, header = TRUE, sep = ",", encoding = 'UTF-8')
can_t <- read.csv(text=x4, header = TRUE, sep = ",", encoding = 'UTF-8')

`%>%` <- magrittr::`%>%`
#------------------ canada data formating ------------------
#format dates;
can_c$date_report<-as.Date(can_c$date_report,format="%d-%m-%y")
can_d$date_death_report<-as.Date(can_d$date_death_report,format="%d-%m-%y")
can_r$date_recovered<-as.Date(can_r$date_recovered,format="%d-%m-%y")
can_t$date_testing<-as.Date(can_t$date_testing,format="%d-%m-%y")

#format province labels;
province_labelc<-c("Alberta","British Columbia","Manitoba","New Brunswick", "Newfoundland and Labrador", "Nova Scotia","Nunavut","NorthWest", "Ontario","Prince Edward Island","Quebec", "Repatriated","Saskatchewan","Yukon")
province_labeld<-c("Alberta","British Columbia","Manitoba","New Brunswick", "Newfoundland and Labrador", "Nova Scotia","Nunavut","NorthWest", "Ontario","Prince Edward Island","Quebec","Saskatchewan","Yukon")

levels(can_c$province)<-province_labelc
levels(can_d$province)<-province_labeld
levels(can_r$province)<-province_labeld
levels(can_t$province)<-province_labeld

#aggregate counts by date combining all province;
can_c_daily <- can_c  %>% group_by(date_report) %>% summarise(c_daily=sum(cases, na.rm = T)) 
can_d_daily <- can_d  %>% group_by(date_death_report) %>% summarise(d_daily=sum(deaths, na.rm = T)) 
can_r_daily <- can_r  %>% group_by(date_recovered) %>% summarise(r_cum=sum(cumulative_recovered, na.rm = T)) 
can_t_daily <- can_t  %>% group_by(can_t$date_testing) %>% summarise(t_cum=sum(cumulative_testing, na.rm=T)) 

#merge all data by dates in a canada dataset;
#this data can be shared on the dashboard; #we can also share the canada data by province;
can<-merge(can_c_daily, can_d_daily, by.x = "date_report",by.y="date_death_report",all.x = T)
can<-merge(can, can_r_daily, by.x = "date_report",by.y="date_recovered",all.x = T)
can<-merge(can, can_t_daily, by.x = "date_report", by.y="can_t$date_testing",all.x = T)
can[is.na(can)]<-0  #give value zero for missing;

can <- can %>% mutate(c_cum = cumsum(c_daily)) #cumulative level;
can <- can %>% mutate(d_cum = cumsum(d_daily)) #cumulative level;
can$a_cum<-can$c_cum - can$r_cum - can$d_cum


#getting daily recovered;
can$r_lag<-lag(can$r_cum)
can$r_lag[is.na(can$r_lag)]<-0
can$r_daily<-can$r_cum-can$r_lag

#first date when reached 100 cases; use this to plot for trajectories;
# can_date_100<-min(can$date_report[can$c_cum>=100],na.rm = T)



`%>%` <- magrittr::`%>%`
#------------World data this is for the trajectory plot------------

x5<-getURL("https://raw.githubusercontent.com/RamiKrispin/coronavirus-csv/master/coronavirus_dataset.csv")
world <- read.csv(text=x5, header = TRUE, sep = ",", encoding = 'UTF-8')

df_us <- world %>% dplyr::filter(type == "confirmed", Country.Region == "US") %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(cases = sum(cases)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(date) %>%
  dplyr::mutate(us = cumsum(cases)) %>%
  dplyr::filter(us > 100)  %>%
  dplyr::select(-cases, -date)

df_us$index <- 1:nrow(df_us)


df_uk <- world %>% dplyr::filter(type == "confirmed", Country.Region == "United Kingdom") %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(cases = sum(cases)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(date) %>%
  dplyr::mutate(uk = cumsum(cases)) %>%
  dplyr::filter(uk > 100)  %>%
  dplyr::select(-cases, -date)

df_uk$index <- 1:nrow(df_uk)


df_sk <- world %>% dplyr::filter(type == "confirmed", Country.Region == "Korea, South") %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(cases = sum(cases)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(date) %>%
  dplyr::mutate(sko = cumsum(cases)) %>%
  dplyr::filter(sko > 100)  %>%
  dplyr::select(-cases, -date)

df_sk$index <- 1:nrow(df_sk)


`%>%` <- magrittr::`%>%`
#------------Canada data this is for the trajectory plot------------
c_cum<- can$c_cum[can$c_cum>100]
df_can<-data.frame(c_cum[complete.cases(c_cum)], 1:length(c_cum[complete.cases(c_cum)]))
names(df_can)<-c("can","index")

#------------provincial data this is for the trajectory plot------------
#now getting data for provinces, only those with more than 100 cases;

ab_cum<- dplyr::filter(can_c, cumulative_cases > 100 & province=="Alberta")  %>% dplyr::select(cumulative_cases)
ab_cum$index <- 1:nrow(ab_cum)  
names(ab_cum)[1]<-"ab"

bc_cum<- dplyr::filter(can_c, cumulative_cases > 100 & province=="British Columbia")  %>% dplyr::select(cumulative_cases)
bc_cum$index <- 1:nrow(bc_cum)  
names(bc_cum)[1]<-"bc"

on_cum<- dplyr::filter(can_c, cumulative_cases > 100 & province=="Ontario")  %>% dplyr::select(cumulative_cases)
on_cum$index <- 1:nrow(on_cum)  
names(on_cum)[1]<-"on"

qc_cum<- dplyr::filter(can_c, cumulative_cases > 100 & province=="Quebec")  %>% dplyr::select(cumulative_cases)
qc_cum$index <- 1:nrow(qc_cum)  
names(qc_cum)[1]<-"qc"


df_trajectory <- df_sk %>% 
  dplyr::left_join(df_us, by = "index") %>%
  dplyr::left_join(df_uk, by = "index") %>%
  dplyr::left_join(df_can, by = "index") %>%
  dplyr::left_join(on_cum, by = "index") %>%
  dplyr::left_join(qc_cum, by = "index") %>%
  dplyr::left_join(ab_cum, by = "index") %>%
  dplyr::left_join(bc_cum, by = "index") 


`%>%` <- magrittr::`%>%`
#------------Marc Olivier plots data prep ------------
#aggregate counts by date, nolonger need to do my own thing thanks to the new time series data! JP life saving!;
# can_p_daily <- can_c  %>% group_by(date_report, province) %>% summarise(c_daily=n()) #individual level;
# can_dp_daily <- can_d  %>% group_by(date_death_report, province) %>% summarise(d_daily=n()) #individual level;

can_p<-merge(can_c, can_d[,c("province","date_death_report","deaths")], 
             by.x = c("date_report","province"),
             by.y=c("date_death_report","province"), all.x = T)

can_p[is.na(can_p)]<-0 #only do this on daily value;
# can_p <- can_p  %>% group_by(province) %>% mutate(c_cum = cumsum(c_daily)) #cumulative level;
can_p <- can_p  %>% group_by(province) %>% mutate(d_cum = cumsum(deaths)) #cumulative level;

can_p<-merge(can_p, can_r, 
             by.x = c("date_report","province"),
             by.y=c("date_recovered","province"), all.x = T)

can_p<-merge(can_p, can_t[,c("date_testing","province","cumulative_testing")], 
             by.x = c("date_report","province"),
             by.y=c("date_testing","province"), all.x = T)

for (i in 7:8){ if (is.na(can_p[1,i])==TRUE) {can_p[1,i]<-0}}
for (j in 7:8){can_p[,j]<- zoo::na.locf(can_p[,j])}

names(can_p)[3:8]<-c("c_daily","c_cum","d_daily","d_cum","r_cum","t_cum")


#------------------ Parameters colours  ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color

#value box colour;
tested_color <- "#00008B"
positive_color <- "#DC143C"
active_color <- "#1E90FF"
recovered_color <- "forestgreen"
death_color <- "#FF0000"
ab_color<-"#000000"
bc_color<-"#E69F00"
mb_color<- "#800000"
nb_color<- "#CC79A7"
nl_color<-"#56B4E9"
ns_color<-"#808000"
nt_color<-"#625D5D"
nu_color<-"#1589FF"
on_color<-"#009E73"
pe_color<-"#0000A0"
qc_color<-"#0072B2"
sk_color<- #800080"
yt_color<-"#E56717"

#canada we will use Red;
us_color<-"deepskyblue" 
uk_color<-"springgreen" 
sko_color<-"violet" 


#------------------ creating Ontario data;  ------------------

on<-can_p[can_p$province=="Ontario",]
on$a_cum<-on$c_cum - on$r_cum - on$d_cum

#getting daily recovered;
on$r_lag<-lag(on$r_cum)
on$r_lag[is.na(on$r_lag)]<-0
on$r_daily<-on$r_cum-on$r_lag

on$c_lag<-lag(on$c_cum)
on$d_lag<-lag(on$d_cum)
on$c_lag[is.na(on$c_lag)]<-0
on$d_lag[is.na(on$d_lag)]<-0

on$c_percentchange<-round(on$c_daily/on$c_lag*100,1)
on$d_percentchange<-round(on$d_daily/on$d_lag*100,1)
on$r_percentchange<-round(on$r_daily/on$r_lag*100,1)

on <- on %>% 
  dplyr::mutate(c_daily_smooth = (c_daily +
                                    dplyr::lag(c_daily, n = 1) +
                                    dplyr::lag(c_daily, n = 2) +
                                    dplyr::lag(c_daily, n = 3) +
                                    dplyr::lag(c_daily, n = 4)) / 5) %>%
  dplyr::mutate(d_daily_smooth = (d_daily +
                                    dplyr::lag(d_daily, n = 1) +
                                    dplyr::lag(d_daily, n = 2) +
                                    dplyr::lag(d_daily, n = 3) +
                                    dplyr::lag(d_daily, n = 4)) / 5) %>%
  dplyr::mutate(r_daily_smooth = (r_daily +
                                    dplyr::lag(r_daily, n = 1) +
                                    dplyr::lag(r_daily, n = 2) +
                                    dplyr::lag(r_daily, n = 3) +
                                    dplyr::lag(r_daily, n = 4)) / 5)

# 0. Health region data - all of this is for the heatmaps;
x6<-getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_hr/cases_timeseries_hr.csv")
x7<-getURL("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_hr/mortality_timeseries_hr.csv")

hr_c <- read.csv(text=x6, header = TRUE, sep = ",", encoding = 'UTF-8')
hr_d <- read.csv(text=x7, header = TRUE, sep = ",", encoding = 'UTF-8')

hr_c$date_report<-as.Date(hr_c$date_report,format="%d-%m-%y")
hr_d$date_death_report<-as.Date(hr_d$date_death_report,format="%d-%m-%y")

# 1. Cases
on_case<- filter(hr_c, province == "Ontario" ) 

on_case_region<- filter(hr_c, province == "Ontario" )  %>% group_by(health_region) %>% summarise(Freq=sum(cases,na.rm = T))
region_order <- unlist(on_case_region[order(on_case_region$Freq),][,1]) #save new region order by total case

on_case$health_region<-factor(on_case$health_region, levels = region_order[1:nrow(on_case_region)])

# 2. deaths
on_death<- filter(hr_d, province == "Ontario" ) 

on_death_region<- filter(hr_d, province == "Ontario" )  %>% group_by(health_region) %>% summarise(Freq=sum(deaths,na.rm = T))
region_order <- unlist(on_death_region[order(on_death_region$Freq),][,1]) #save new region order by total case

on_death$health_region<-factor(on_death$health_region, levels = region_order[1:nrow(on_case_region)])


#------------------ creating Toronto data;  ------------------

#aggregate counts by date;
trt_c_daily <- filter(hr_c, health_region=="Toronto")  %>% group_by(date_report) %>% select(-cumulative_cases, -province)
trt_d_daily <- filter(hr_d, health_region=="Toronto")  %>% group_by(date_death_report) %>% select(-cumulative_deaths, -province)

#this data can be shared on the dashboard; #we can also share the canada data by province;
trt<-merge(trt_c_daily, trt_d_daily, by.x = c("date_report", "health_region"),by.y=c("date_death_report","health_region"),all.x = T)

trt[is.na(trt)]<-0  #give value zero for missing; only do this on daily value!
names(trt)[3:4]<-c("c_daily","d_daily")

trt <- trt %>% mutate(c_cum = cumsum(c_daily)) #cumulative level;
trt <- trt %>% mutate(d_cum = cumsum(d_daily)) #cumulative level;
trt$a_cum <- trt$c_cum - trt$d_cum

df_trt<-data.frame(trt$c_cum[trt$c_cum>100], c(1:length(trt$c_cum[trt$c_cum>100])))
names(df_trt)<-c("trt","index")

df_trajectory <- df_trajectory %>% 
  dplyr::left_join(df_trt, by = "index")
 
#------------------ creating quebec data;  ------------------

qc<-can_p[can_p$province=="Quebec",]
qc$a_cum<-qc$c_cum - qc$r_cum - qc$d_cum


#getting daily recovered;
qc$r_lag<-lag(qc$r_cum)
qc$r_lag[is.na(qc$r_lag)]<-0
qc$r_daily<-qc$r_cum-qc$r_lag

qc$c_lag<-lag(qc$c_cum)
qc$d_lag<-lag(qc$d_cum)
qc$c_lag[is.na(qc$c_lag)]<-0
qc$d_lag[is.na(qc$d_lag)]<-0

qc$c_percentchange<-round(qc$c_daily/qc$c_lag*100,1)
qc$d_percentchange<-round(qc$d_daily/qc$d_lag*100,1)
qc$r_percentchange<-round(qc$r_daily/qc$r_lag*100,1)

qc <- qc %>% 
  dplyr::mutate(c_daily_smooth = (c_daily +
                                    dplyr::lag(c_daily, n = 1) +
                                    dplyr::lag(c_daily, n = 2) +
                                    dplyr::lag(c_daily, n = 3) +
                                    dplyr::lag(c_daily, n = 4)) / 5) %>%
  dplyr::mutate(d_daily_smooth = (d_daily +
                                    dplyr::lag(d_daily, n = 1) +
                                    dplyr::lag(d_daily, n = 2) +
                                    dplyr::lag(d_daily, n = 3) +
                                    dplyr::lag(d_daily, n = 4)) / 5) %>%
  dplyr::mutate(r_daily_smooth = (r_daily +
                                    dplyr::lag(r_daily, n = 1) +
                                    dplyr::lag(r_daily, n = 2) +
                                    dplyr::lag(r_daily, n = 3) +
                                    dplyr::lag(r_daily, n = 4)) / 5)


# 1. Cases
qc_case<- filter(hr_c, province == "Quebec" ) 

qc_case_region<- filter(hr_c, province == "Quebec" )  %>% group_by(health_region) %>% summarise(Freq=sum(cases,na.rm = T))
region_order <- unlist(qc_case_region[order(qc_case_region$Freq),][,1]) #save new region order by total case

qc_case$health_region<-factor(qc_case$health_region, levels = region_order[1:nrow(qc_case_region)])

# 2. deaths
qc_death<- filter(hr_d, province == "Quebec" ) 

qc_death_region<- filter(hr_d, province == "Quebec" )  %>% group_by(health_region) %>% summarise(Freq=sum(deaths,na.rm = T))
region_order <- unlist(qc_death_region[order(qc_death_region$Freq),][,1]) #save new region order by total case

qc_death$health_region<-factor(qc_death$health_region, levels = region_order[1:nrow(qc_case_region)])


```



Summary {data-orientation=rows}
=======================================================================

Row
-----------------------------------------------------------------------
### tested {.value-box}
```{r}
valueBox(value = paste(format(max(can$t_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Tested Cases", 
         icon = "fas fa-user-md", 
         color = tested_color)
```



### Positive Cases {.value-box}

```{r}
valueBox(value = paste(format(max(can$c_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Positive", 
         icon = "far fa-plus-square", 
         color = positive_color)
```


### Active {.value-box}
```{r}
valueBox(value = paste(format(max(can$a_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Active Cases", 
         icon = "fas fa-ambulance",
         color = active_color)
```

### recovered {.value-box}
```{r}
valueBox(value = paste(format(max(can$r_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Recovered Cases", 
         icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### death {.value-box}
```{r}
valueBox(value = paste(format(max(can$d_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Deaths", 
         icon = "", 
         color = death_color)
```

Row
-----------------------------------------------------------------------

### Overall Distribution of Cases

```{r}
plotly::plot_ly(data = can[can$date_report > as.Date("2020-02-29"),],
        x = ~ date_report,
        y = ~ a_cum, 
        name = 'Active', 
        fillcolor = '#1f77b4',
        type = 'scatter',
        mode = 'none', 
        stackgroup = 'one') %>%
  plotly::add_trace( y = ~ d_cum, 
             name = "Death",
             fillcolor = '#E41317') %>%
  plotly::add_trace(y = ~ r_cum, 
            name = 'Recovered', 
            fillcolor = 'forestgreen') %>%
  plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Number of Cumulative Cases"),
         hovermode = "compared")
```

### Trajectory of Confirmed Cases in Canada by Province (with more than 1000 cases)

```{r}
# `%>%` <- magrittr::`%>%`
plotly::plot_ly(data = df_trajectory) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ us,
                    name = "United States",  line = list(width = 2,color = us_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can,
                    line = list(color = "red", width = 4),
                    name = "Canada") %>%
  plotly::add_lines(x = ~ index,
                    y = ~ uk,
                    name = "United Kingdom",  line = list(width = 2,color = uk_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ sko,
                    name = "South Korea",  line = list(color = sko_color, width = 2)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ qc,
                    name = "Quebec",  line = list(width = 2, color=qc_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ on,
                    name = "Ontario", line = list(width = 2, color=on_color)) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ ab,
                    name = "Alberta",  line = list(width = 2, color=ab_color)) %>%   
  plotly::add_lines(x = ~ index,
                    y = ~ bc,
                    name = "British Columbia",  line = list(width = 2, color=bc_color)) %>%
  plotly::layout(yaxis = list(title = "Cumulative Positive Cases (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total positive cases surpass 100"))

                 # legend = list(x = 0.7, y = 0.1)

```


Row
-----------------------------------------------------------------------

### Distribution of Daily New Confirmed, Fatal and Recovered Cases

```{r}
can <- can %>% 
  dplyr::mutate(c_daily_smooth = (c_daily +
                                    dplyr::lag(c_daily, n = 1) +
                                    dplyr::lag(c_daily, n = 2) +
                                    dplyr::lag(c_daily, n = 3) +
                                    dplyr::lag(c_daily, n = 4)) / 5) %>%
  dplyr::mutate(d_daily_smooth = (d_daily +
                                    dplyr::lag(d_daily, n = 1) +
                                    dplyr::lag(d_daily, n = 2) +
                                    dplyr::lag(d_daily, n = 3) +
                                    dplyr::lag(d_daily, n = 4)) / 5) %>%
  dplyr::mutate(r_daily_smooth = (r_daily +
                                    dplyr::lag(r_daily, n = 1) +
                                    dplyr::lag(r_daily, n = 2) +
                                    dplyr::lag(r_daily, n = 3) +
                                    dplyr::lag(r_daily, n = 4)) / 5)

fig <- plot_ly(can[can$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_daily, type="scatter", mode = 'markers',  name = 'Daily New Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(x = ~ date_report, y = ~ c_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Case Trend', line = list(color = '#1f77b4',width=2), marker = list(color = '#1f77b4', opacity=0)) %>% 
  add_trace(y = ~d_daily, name = 'Daily New Death', marker = list(color = 'red'), mode = 'markers') %>% 
  add_trace(x = ~ date_report, y = ~ d_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Fatal Trend', line = list(color = 'red',width=2), marker = list(color = 'red', opacity=0)) %>% 
  add_trace(y = ~r_daily, name = 'Daily New Recovered', marker = list(color = 'forestgreen'), mode = 'markers') %>% 
  add_trace(y = ~r_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Recovery Trend',line = list(color = 'forestgreen',width=2), marker = list(color = 'forestgreen', opacity=0)) 


fig %>% plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = "Using 5 days trailing moving average to calculate the trend lines"),
         yaxis = list(title = "Number of New Daily Cases"))

```


### Daily Percentage Increase on Confirmed and Fatal Cases


```{r}
can$c_lag<-lag(can$c_cum)
can$d_lag<-lag(can$d_cum)
can$c_lag[is.na(can$c_lag)]<-0
can$d_lag[is.na(can$d_lag)]<-0

can$c_percentchange<-round(can$c_daily/can$c_lag*100,1)
can$d_percentchange<-round(can$d_daily/can$d_lag*100, 1)
can$r_percentchange<-round(can$r_daily/can$r_lag*100,1)


fig <- plot_ly(can[can$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_percentchange, type="scatter", mode = 'lines+markers',  name = '% Change Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(y = ~d_percentchange, name = '% Change Death', marker = list(color = 'red'), line=list(color="red"),mode = 'lines+markers') 
# %>%  add_trace(y = ~r_percentchange, name = '% Change Recovered', marker = list(color = 'forestgreen'), mode = 'lines+markers')

fig %>% plotly::layout(title = "",
         xaxis = list(title = ""),
         legend = list(x = 0.7, y = 0.95),
         yaxis = list(title = "Percentage change in Cases"))


```


Row
-----------------------------------------------------------------------

### Marc's Distribution of Confirmed Cases by Province and Territory

```{r}
can_c_mo<- can_p[can_p$date_report>as.Date("2020-02-29"),c("date_report","province","c_cum")]
can_c_mo<- reshape(can_c_mo, idvar = "date_report",timevar="province",direction = "wide")

names(can_c_mo)[2:ncol(can_c_mo)]<-sub("c.cum.", "", names(can_c_mo)[2:ncol(can_c_mo)])
can_c_mo[is.na(can_c_mo)]<-0

fig <- plot_ly(can_c_mo, x = ~date_report, y = ~Quebec, name = 'Quebec', type = 'scatter', mode = 'none', stackgroup = 'one', groupnorm = 'percent', fillcolor = qc_color)
fig <- fig %>% add_trace(y = ~Ontario, name = 'Ontario', fillcolor = on_color)
fig <- fig %>% add_trace(y = ~`British Columbia`, name = 'British Columbia', fillcolor = bc_color)
fig <- fig %>% add_trace(y = ~Alberta, name = 'Alberta', fillcolor = ab_color)
fig <- fig %>% add_trace(y = ~`Nova Scotia`, name = 'Nova Scotia', fillcolor = ns_color)
fig <- fig %>% add_trace(y = ~Saskatchewan, name = 'Saskatchewan', fillcolor = sk_color)
fig <- fig %>% add_trace(y = ~Manitoba, name = 'Manitoba', fillcolor = mb_color)
fig <- fig %>% add_trace(y = ~`Newfoundland and Labrador`, name = 'Newfoundland and Labrador', fillcolor = nl_color)
fig <- fig %>% add_trace(y = ~`New Brunswick`, name = 'New Brunswick', fillcolor = nb_color)
fig <- fig %>% add_trace(y = ~`Prince Edward Island`, name = 'Prince Edward Island', fillcolor = pe_color)
fig <- fig %>% add_trace(y = ~Repatriated, name = 'Repatriated', fillcolor = "lightgrey")
fig <- fig %>% add_trace(y = ~Yukon, name = 'Yukon', fillcolor = yt_color)
fig <- fig %>% add_trace(y = ~NorthWest, name = 'NorthWest', fillcolor = nt_color)
fig <- fig %>% layout(
         xaxis = list(title = "",
                      showgrid = FALSE),
         yaxis = list(title = "Proportion from the Total",
                      showgrid = FALSE,
                      ticksuffix = '%'))
fig

```


### Marc's Distribution of Fatal Cases by Province and Territory

```{r}
can_d_mo<- can_p[can_p$date_report>as.Date("2020-03-07"),c("date_report","province","d_cum")]
can_d_mo<- reshape(can_d_mo, idvar = "date_report",timevar="province",direction = "wide")

names(can_d_mo)[2:ncol(can_d_mo)]<-sub("d.cum.", "", names(can_d_mo)[2:ncol(can_d_mo)])
can_d_mo[is.na(can_d_mo)]<-0

fig <- plot_ly(can_d_mo, x = ~date_report, y = ~Quebec, name = 'Quebec', type = 'scatter', mode = 'none', stackgroup = 'one', groupnorm = 'percent', fillcolor = qc_color)
fig <- fig %>% add_trace(y = ~Ontario, name = 'Ontario', fillcolor = on_color)
fig <- fig %>% add_trace(y = ~`British Columbia`, name = 'British Columbia', fillcolor = bc_color)
fig <- fig %>% add_trace(y = ~Alberta, name = 'Alberta', fillcolor = ab_color)
fig <- fig %>% add_trace(y = ~`Nova Scotia`, name = 'Nova Scotia', fillcolor = ns_color)
fig <- fig %>% add_trace(y = ~Saskatchewan, name = 'Saskatchewan', fillcolor = sk_color)
fig <- fig %>% add_trace(y = ~Manitoba, name = 'Manitoba', fillcolor = mb_color)
fig <- fig %>% add_trace(y = ~`Newfoundland and Labrador`, name = 'Newfoundland and Labrador', fillcolor = nl_color)
fig <- fig %>% add_trace(y = ~`New Brunswick`, name = 'New Brunswick', fillcolor = nb_color)
fig <- fig %>% add_trace(y = ~`Prince Edward Island`, name = 'Prince Edward Island', fillcolor = pe_color)
fig <- fig %>% add_trace(y = ~Repatriated, name = 'Repatriated', fillcolor = "lightgrey")
fig <- fig %>% add_trace(y = ~Yukon, name = 'Yukon', fillcolor = yt_color)
fig <- fig %>% add_trace(y = ~NorthWest, name = 'NorthWest', fillcolor = nt_color)
fig <- fig %>% layout(
         xaxis = list(title = "",
                      showgrid = FALSE),
         yaxis = list(title = "Proportion from the Total",
                      showgrid = FALSE,
                      ticksuffix = '%'))
fig

```


Toronto {data-orientation=rows}
=======================================================================

Row
-----------------------------------------------------------------------


### Positive Cases {.value-box}

```{r}
valueBox(value = paste(format(max(trt$c_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Positive", 
         icon = "far fa-plus-square", 
         color = positive_color)
```

### Active and Recovered Cases {.value-box}
```{r}
valueBox(value = paste(format(max(trt$a_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Active and Recovered Cases", 
         icon = "fas fa-ambulance",
         color = active_color)
```


### death {.value-box}
```{r}
valueBox(value = paste(format(max(trt$d_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Deaths", 
         icon = "", 
         color = death_color)
```

Row
-----------------------------------------------------------------------

### Overall Distribution of Cases

```{r}
plotly::plot_ly(data = trt[trt$date_report>as.Date("2020-02-29"),],
        x = ~ date_report,
        y = ~ a_cum, 
        name = 'Active and Recovered Cases', 
        fillcolor = '#1f77b4',
        type = 'scatter',
        mode = 'none', 
        stackgroup = 'one') %>%
  plotly::add_trace( y = ~ d_cum, 
             name = "Death",
             fillcolor = '#E41317') %>%
  plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Number of Cumulative Cases"),
         hovermode = "compared")
```


### Distribution of Daily New Confirmed, Fatal and Recovered Cases

```{r}
trt <- trt %>% 
  dplyr::mutate(c_daily_smooth = (c_daily +
                                    dplyr::lag(c_daily, n = 1) +
                                    dplyr::lag(c_daily, n = 2) +
                                    dplyr::lag(c_daily, n = 3) +
                                    dplyr::lag(c_daily, n = 4)) / 5) %>%
  dplyr::mutate(d_daily_smooth = (d_daily +
                                    dplyr::lag(d_daily, n = 1) +
                                    dplyr::lag(d_daily, n = 2) +
                                    dplyr::lag(d_daily, n = 3) +
                                    dplyr::lag(d_daily, n = 4)) / 5) 

fig <- plot_ly(trt[trt$date_report>as.Date("2020-02-29"),], x = ~ date_report, y = ~c_daily, type="scatter", mode = 'markers',  name = 'Daily New Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(x = ~ date_report, y = ~ c_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Case Trend', line = list(color = '#1f77b4',width=2), marker = list(color = '#1f77b4', opacity=0)) %>% 
  add_trace(y = ~d_daily, name = 'Daily New Death', marker = list(color = 'red'), mode = 'markers') %>% 
  add_trace(x = ~ date_report, y = ~ d_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Fatal Trend', line = list(color = 'red',width=2), marker = list(color = 'red', opacity=0)) 


fig %>% plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = "Using 5 days trailing moving average to calculate the trend lines"),
         yaxis = list(title = "Number of New Daily Cases"))

```

### Daily Percentage Increase on Confirmed and Fatal Cases

```{r}
trt$c_lag<-lag(trt$c_cum)
trt$d_lag<-lag(trt$d_cum)
trt$c_lag[is.na(trt$c_lag)]<-0
trt$d_lag[is.na(trt$d_lag)]<-0

trt$c_percentchange<-round(trt$c_daily/trt$c_lag*100,1)
trt$d_percentchange<-round(trt$d_daily/trt$d_lag*100,1)


fig <- plot_ly(trt[trt$date_report>as.Date("2020-02-29"),], x = ~ date_report, y = ~c_percentchange, type="scatter", mode = 'lines+markers',  name = '% Change Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(y = ~d_percentchange, name = '% Change Death', marker = list(color = 'red'), line=list(color="red"),mode = 'lines+markers') 
# %>%  add_trace(y = ~r_percentchange, name = '% Change Recovered', marker = list(color = 'forestgreen'), mode = 'lines+markers')

fig %>% plotly::layout(title = "",
         xaxis = list(title = ""),
         legend = list(x = 0.7, y = 0.95),
         yaxis = list(title = "Percentage change in Cases"))

```


Row
-----------------------------------------------------------------------

### Trajectory of Confirmed Cases in Toronto 

```{r}
# `%>%` <- magrittr::`%>%`
plotly::plot_ly(data = df_trajectory) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ trt,
                    name = "Toronto",  line = list(width = 4,color = "darkblue")) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can,
                    line = list(color = "red", width = 4),
                    name = "Canada") %>%
  plotly::add_lines(x = ~ index,
                    y = ~ us,
                    name = "United States",  line = list(width = 2,color = us_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ uk,
                    name = "United Kingdom",  line = list(width = 2,color = uk_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ sko,
                    name = "South Korea",  line = list(color = sko_color, width = 2)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ qc,
                    name = "Quebec",  line = list(width = 2, color=qc_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ on,
                    name = "Ontario", line = list(width = 2, color=on_color)) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ ab,
                    name = "Alberta",  line = list(width = 2, color=ab_color)) %>%   
  plotly::add_lines(x = ~ index,
                    y = ~ bc,
                    name = "British Columbia",  line = list(width = 2, color=bc_color)) %>%
  plotly::layout(yaxis = list(title = "Cumulative Positive Cases (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total positive cases surpass 100"))

                 # legend = list(x = 0.7, y = 0.1)

```



### Toronto Data

```{r}
trt[,c(1,3:6)] %>% 
  DT::datatable(rownames = FALSE,
                colnames = c("Date","Daily Cases","Daily Deaths", "Cumulative Cases", "Cumulative Deaths"), options = list(searching = FALSE), filter = 'none')

```




Ontario {data-orientation=rows}
=======================================================================

Column {.sidebar data-width=620 data-padding=10}
-------------------------------------
##### Overall Distribution of Cases
    
```{r}
plotly::plot_ly(data = on[on$date_report> as.Date("2020-02-29"),],
        x = ~ date_report,
        y = ~ a_cum, 
        name = 'Active', 
        fillcolor = '#1f77b4',
        type = 'scatter',
        mode = 'none', 
        stackgroup = 'one') %>%
  plotly::add_trace( y = ~ d_cum, 
             name = "Death",
             fillcolor = '#E41317') %>%
  plotly::add_trace(y = ~ r_cum, 
            name = 'Recovered', 
            fillcolor = 'forestgreen') %>%
  plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Number of Cumulative Cases"),
         hovermode = "compared")
```
   

##### Distribution of Daily New Confirmed, Fatal and Recovered Cases

    
```{r}

fig <- plot_ly(on[on$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_daily, type="scatter", mode = 'markers',  name = 'Daily New Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(x = ~ date_report, y = ~ c_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Case Trend', line = list(color = '#1f77b4',width=2), marker = list(color = '#1f77b4', opacity=0)) %>% 
  add_trace(y = ~d_daily, name = 'Daily New Death', marker = list(color = 'red'), mode = 'markers') %>% 
  add_trace(x = ~ date_report, y = ~ d_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Fatal Trend', line = list(color = 'red',width=2), marker = list(color = 'red', opacity=0)) %>% 
  add_trace(y = ~r_daily, name = 'Daily New Recovered', marker = list(color = 'forestgreen'), mode = 'markers') %>% 
  add_trace(y = ~r_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Recovery Trend',line = list(color = 'forestgreen',width=2), marker = list(color = 'forestgreen', opacity=0)) 


fig %>% plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = "Using 5 days trailing moving average to calculate the trend lines"),
         yaxis = list(title = "Number of New Daily Cases"))

```


##### Daily Percentage Increase on Confirmed and Fatal Cases

```{r}
fig <- plot_ly(on[on$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_percentchange, type="scatter", mode = 'lines+markers',  name = '% Change Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(y = ~d_percentchange, name = '% Change Death', marker = list(color = 'red'), line=list(color="red"),mode = 'lines+markers') 
# %>%  add_trace(y = ~r_percentchange, name = '% Change Recovered', marker = list(color = 'forestgreen'), mode = 'lines+markers')

fig %>% plotly::layout(title = "",
         legend = list(x = 0.7, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Percentage change in Cases"))

```


Row 
-------------------------------------
### tested {.value-box}
```{r}
valueBox(value = paste(format(max(on$t_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Tested Cases", 
         icon = "fas fa-user-md", 
         color = tested_color)
```


### Positive Cases {.value-box}

```{r}
valueBox(value = paste(format(max(on$c_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Positive", 
         icon = "far fa-plus-square", 
         color = positive_color)
```


### Active {.value-box}
```{r}
valueBox(value = paste(format(max(on$a_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Active Cases", 
         icon = "fas fa-ambulance",
         color = active_color)
```

### recovered {.value-box}
```{r}
valueBox(value = paste(format(max(on$r_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Recovered Cases", 
         icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### death {.value-box}
```{r}
valueBox(value = paste(format(max(on$d_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Deaths", 
         icon = "", 
         color = death_color)
```


Row {data-height=650 .tabset} 
-------------------------------------
### Heatmaps on Cumulative Cases by Regions

```{r }

plot_ly(on_case[on_case$date_report>as.Date("2020-02-29"),] , 
               x=~date_report, y=~health_region, z = ~ cumulative_cases, type = "heatmap",
               colors = "YlGnBu",zmin=0, zmax=350,
               xgap=1, ygap=1) %>% plotly::layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))


```   

 
### Heatmaps on Daily New Cases by Regions

```{r}
plot_ly(on_case[on_case$date_report>as.Date("2020-02-29"),] , 
               x=~date_report, y=~health_region, z = ~ cases, type = "heatmap",
               colors = "YlGnBu",zmin=0, zmax=50,
               xgap=1, ygap=1) %>% plotly::layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))

```
    


Row {data-height=650 .tabset} 
-------------------------------------

### Heatmaps on Cumulative Deaths by Regions

```{r }

plot_ly(on_death[on_death$date_death_report>as.Date("2020-02-29"),] , 
               x=~date_death_report, y=~health_region, z = ~ cumulative_deaths, type = "heatmap",
               colors = colorRamp(c("white","red")),zmin=0,zmax=50, xgap=1, ygap=1) %>% plotly::layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))



```   

 
### Heatmaps on Daily New Deaths by Regions

```{r}
plot_ly(on_death[on_death$date_death_report>as.Date("2020-02-29"),] , 
               x=~date_death_report, y=~health_region, z = ~ deaths, type = "heatmap",
               colors = colorRamp(c("white","red")),zmin=0,zmax=10, xgap=1, ygap=1) %>% plotly::layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))

```


Quebec
=======================================================================

Column {.sidebar data-width=620 data-padding=10}
-------------------------------------

##### Overall Distribution of Cases

    
```{r}
qc2<-filter(qc, date_report > as.Date("2020-02-29"))
plotly::plot_ly(data = qc2,
        x = ~ date_report,
        y = ~ a_cum, 
        name = 'Active', 
        fillcolor = '#1f77b4',
        type = 'scatter',
        mode = 'none', 
        stackgroup = 'one') %>%
  plotly::add_trace( y = ~ d_cum, 
             name = "Death",
             fillcolor = '#E41317') %>%
  plotly::add_trace(y = ~ r_cum, 
            name = 'Recovered', 
            fillcolor = 'forestgreen') %>%
  plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Number of Cumulative Cases"),
         hovermode = "compared")
```
   

##### Distribution of Daily New Confirmed, Fatal and Recovered Cases

    
```{r}

fig <- plot_ly(qc[qc$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_daily, type="scatter", mode = 'markers',  name = 'Daily New Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(x = ~ date_report, y = ~ c_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Case Trend', line = list(color = '#1f77b4',width=2), marker = list(color = '#1f77b4', opacity=0)) %>% 
  add_trace(y = ~d_daily, name = 'Daily New Death', marker = list(color = 'red'), mode = 'markers') %>% 
  add_trace(x = ~ date_report, y = ~ d_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Fatal Trend', line = list(color = 'red',width=2), marker = list(color = 'red', opacity=0)) %>% 
  add_trace(y = ~r_daily, name = 'Daily New Recovered', marker = list(color = 'forestgreen'), mode = 'markers') %>% 
  add_trace(y = ~r_daily_smooth, type = 'scatter', mode = 'lines+markers', name = 'Daily Recovery Trend',line = list(color = 'forestgreen',width=2), marker = list(color = 'forestgreen', opacity=0)) 


fig %>% plotly::layout(title = "",
         legend = list(x = 0.05, y = 0.95),
         xaxis = list(title = "Using 5 days trailing moving average to calculate the trend lines"),
         yaxis = list(title = "Number of New Daily Cases"))

```

##### Daily Percentage Increase on Confirmed and Fatal Cases
    
```{r}
fig <- plot_ly(qc[qc$date_report>as.Date("2020-03-01"),], x = ~ date_report, y = ~c_percentchange, type="scatter", mode = 'lines+markers',  name = '% Change Case', marker = list(color = '#1f77b4'))  %>% 
  add_trace(y = ~d_percentchange, name = '% Change Death', marker = list(color = 'red'), line=list(color="red"),mode = 'lines+markers') 
# %>%  add_trace(y = ~r_percentchange, name = '% Change Recovered', marker = list(color = 'forestgreen'), mode = 'lines+markers')

fig %>% plotly::layout(title = "",
         legend = list(x = 0.7, y = 0.95),
         xaxis = list(title = ""),
         yaxis = list(title = "Percentage change in Cases"))

```

Row 
-------------------------------------
### tested {.value-box}
```{r}
valueBox(value = paste(format(max(qc$t_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Tested Cases", 
         icon = "fas fa-user-md", 
         color = tested_color)
```


### Positive Cases {.value-box}

```{r}
valueBox(value = paste(format(max(qc$c_cum, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Positive", 
         icon = "far fa-plus-square", 
         color = positive_color)
```


### Active {.value-box}
```{r}
valueBox(value = paste(format(max(qc$a_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Active Cases", 
         icon = "fas fa-ambulance",
         color = active_color)
```

### recovered {.value-box}
```{r}
valueBox(value = paste(format(max(qc$r_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Recovered Cases", 
         icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### death {.value-box}
```{r}
valueBox(value = paste(format(max(qc$d_cum, na.rm = T), big.mark = ","), sep = ""),
         caption = "Deaths", 
         icon = "", 
         color = death_color)
```


Row {data-height=650 .tabset} 
-------------------------------------
    
### Heatmaps on Cumulative Cases by Regions

```{r }
plot_ly(qc_case[qc_case$date_report>as.Date("2020-02-29"),], x=~date_report, y=~health_region, z = ~ cumulative_cases, type = "heatmap", colors = "YlGnBu",zmin=0, zmax=350,
               xgap=1, ygap=1) %>% layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))
```   

 
### Heatmaps on Daily New Cases by Regions

```{r}
plot_ly(qc_case[qc_case$date_report>as.Date("2020-02-29"),], x=~date_report, y=~health_region, z = ~ cases, type = "heatmap", colors = "YlGnBu",zmin=0, zmax=50,
               xgap=1, ygap=1) %>% layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))

```


Row {data-height=650 .tabset} 
-------------------------------------

### Heatmaps on Cumulative Deaths by Regions

```{r }
plot_ly(qc_death[qc_death$date_death_report>as.Date("2020-02-29"),] , 
               x=~date_death_report, y=~health_region, z = ~ cumulative_deaths, type = "heatmap",
               colors = colorRamp(c("white","red")),zmin=0,zmax=50, xgap=1, ygap=1) %>% layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))
```   

 
### Heatmaps on Daily New Deaths by Regions

```{r}
plot_ly(qc_death[qc_death$date_death_report>as.Date("2020-02-29"),] , 
               x=~date_death_report, y=~health_region, z = ~ deaths, type = "heatmap",
               colors = colorRamp(c("white","red")),zmin=0,zmax=10, xgap=1, ygap=1) %>% layout(showlegend=FALSE,title = "",
         xaxis = list(title = "", side="top"),
         yaxis = list(title = ""))

```


Canada Data
=======================================================================

```{r}
can[,c(1,2,3,10,6,7,4)] %>% 
  DT::datatable(rownames = FALSE,
                colnames = c("Date","Daily New Cases","Daily New Death","Daily New Recovered","Cumulative Cases", "Cumulative Death", "Cumulative Recovered"), options = list(searchHighlight = TRUE, 
                           pageLength = nrow(can)), filter = 'top')
```


About
=======================================================================
**Time of last update** `r pb.date <- as.POSIXct(Sys.time()); format(pb.date, tz="America/Toronto",usetz=TRUE)`

**From the author**

Thank you for your interest in my new data visualization project. There are many great COVID-19 Canada dashboards available online. I have compiled the following list to show some of the dashboards that inspired me to build this one. 

- [The COVID-19 Canada Open Data Working Groups dashboard](https://art-bd.shinyapps.io/covid19canada/) made by [Jean-Paul R. Soucy](https://twitter.com/JPSoucy) and [Isha Berry](https://twitter.com/ishaberry2), two amazing talented students from my school, DLSPH, at Univerity of Toronto.

- [The Covid19 Italy dashboard](https://github.com/RamiKrispin/italy_dash) by [Rami Krispin](https://twitter.com/Rami_Krispin). I used Rami's [R markdown code](https://github.com/RamiKrispin/covid19Italy) for this dashboard.

- [The Esri Canada COVID-19 Canada dashboard](https://resources-covid19canada.hub.arcgis.com/).

- [The PHAC COVID-19 Canada dashboard](https://experience.arcgis.com/experience/2f1a13ca0b29422f9b34660f0b705043/), **most accurate and reliable** info on Canadian COVID-19 epidemiology data.

- [The Ontario-specific COVID-19 Analysis Website](https://howsmyflattening.ca/#/analysis) which covers detailed Ontario focused analytical information on critical care analysis, health region analysis and testing analysis. This website I believe was also built by a group of University of Toronto students.

- [The Québec provincial COVID-19 dashboard](https://www.inspq.qc.ca/covid-19/donnees).

- [The Alberta provincial COVID-19 dashboard](https://covid19stats.alberta.ca/).
 
Using the four COVID-19 Canada Open Data Working Group's spreadsheets on cases, mortality, recovery and testing at <https://github.com/ishaberry/Covid19Canada>, I created an aggregated Canada COVID-19 daily data. You can view this data on the data page of this dash.


Last but not least, I want to give a shout-out to [Marc-Olivier Hétu](https://twitter.com/suivicovid) whom I met on [twitter](https://twitter.com/suivicovid). He has been scraping public COVID-19 Canada data since March and compiled an outstanding spreadsheet with graphic analysis, which he generiously shared with me, because he said he's glad that more people get to use his data to build tools. I have adopted one of Marc's special figure on this dash and I named it the Marc-Olivier figure to carry on the legacy.



**Acknowledgement**

 - **Data** 
    - Canada data used in this dashboard are extracted from the data speardsheet posted by the COVID-19 Canada Open Data Working Group (https://github.com/ishaberry/Covid19Canada) lead by Jean-Paul R. Soucy and Isha Berry.
    - US data used in this dashboard are from the R package coronavirus developed by [Rami Krispin](https://github.com/RamiKrispin/coronavirus) which is generated from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [website](https://systems.jhu.edu/research/public-health/ncov/).
 - **Dashboard R code**
    - I built this dashboard using Rami Krispin's Covid19 Italy [dashboard](https://github.com/RamiKrispin/italy_dash)'s code as template. Rami's code can be found on this [github page](https://github.com/RamiKrispin/covid19Italy).
 - **Special thanks** to my supervisor Dr. [Eleanor Pullenayegum](https://twitter.com/EMPullenayegum) at [Sickkids](http://www.sickkids.ca/AboutSickKids/Directory/People/P/Eleanor-Pullenayegum-staff-profile.html) for her support on all my COVID-19 projects!
 
**I would not have made this bashboard if not for the work of the above team and individuals.** You can find the code of this dashboard at <https://github.com/Kuan-Liu/canada_dash>.


**Feedbacks**

Please feel free to reach me at kuan.liu@mail.utoronto.ca. You can also find me on [Twitter](https://twitter.com/KuanLiu2).


**Message**

**With a budget shinyapps.io membership, there is a restriction on the number of hours our dashboard can stay active without going over the current allowance.** 

**Please help us stay within our plan by closing the dashboard webpage once you have completed the viewing.** 

**If you would like to get in touch about hosting the dash at your webpage with shiny server or Rstudio connect, please get in touch with us.**
    
